Handling Recursion in Generic Programming Using Closed Type Families

### 2018-08-22 lncs-paper.tex
Change to the abstract description

subtermsNP :: ∀a xs. NP I xs -> [a]
subtermsNP p (I y :* ys)
  | typeOf @a y = castEq y : subtermsNP ys
  | otherwise   = subtermsNP ys
subtermsNP _ Nil = []
\end{lstlisting}
The function \K{subterms} translates the term to its representation, unwrapping the sum of products from \K{SOP}, and passes that to the auxiliary function \K{subtermsNS}. The latter merely traverses the sum and, once reaches the product,  passes it further to \K{subtermsNP}.

@@

The low-level definition of \K{subtermsNP} using the type cast resembles its outline given in the previous section:
\begin{lstlisting}
subtermsNP :: ∀a xs. All (ProofEq a) xs => NP I xs -> [a]
subtermsNP (I (y :: x) :* ys) =
  case castEq y of
    Just t  -> t : subtermsNP ys
    Nothing -> subtermsNP  ys
subtermsNP _ Nil = []
\end{lstlisting}

@@

instance Show  b => CaseShow 'False a b where
  caseShow' = show
instance GShow a => CaseShow 'True  a a where
  caseShow' = gshow
\end{lstlisting}
We provide a synonym for the \K{CaseShow (Equal a b) a b} instance, which we call \K{CaseRecShow}, as before with \K{ProofEq}; likewise a synonym for the matching function:
\begin{lstlisting}
caseShow :: ∀a b. CaseRecShow a b => b -> String
caseShow t = caseShow' @(Equal a b) @a t
\end{lstlisting}

@@

We can continue along the lines of the previous examples to consider the approach separately from concrete ones. One can use the technique abstractedly, following the pattern below:

\begin{lstlisting}
class DispatchRec_Aux (p :: Bool)
                    (a_1 :: k_1) ... (a_n :: k_n) (b :: *) where
  dispatchRec_Aux :: b -> X

@@

dispatchRec :: ∀a_1 ... a_n b. DispatchRec a_1 ... a_n b => b -> X
dispatchRec = dispatchRec_Aux @(P a_1 ... a_k b) @a_1 ... @a_k ... @a_n
\end{lstlisting}

@@

>>>>

subtermsNP :: ∀a xs. NP I xs -> [a]
subtermsNP (I y :* ys)
  | typeOf @a y = castEq y : subtermsNP ys
  | otherwise   = subtermsNP ys
subtermsNP Nil = []
\end{lstlisting}
The function \K{subterms} translates the term to its representation, unwrapping the sum of products from \K{SOP}, and passes that to the auxiliary function \K{subtermsNS}. The latter merely traverses the sum and, once reaches the product,  passes it further to \K{subtermsNP}.

@@

The low-level definition of \K{subtermsNP} using the type cast resembles its outline given in the previous section:
\begin{lstlisting}
subtermsNP :: ∀a xs. All (ProofEq a) xs => NP I xs -> [a]
subtermsNP (I (y :: x) :* ys) =
  case castEq y of
    Just t  -> t : subtermsNP ys
    Nothing -> subtermsNP  ys
subtermsNP Nil = []
\end{lstlisting}

@@

instance Show  b => CaseShow 'False a b where
  caseShow' = show
instance GShow a => CaseShow 'True  a a where
  caseShow' = gshow
\end{lstlisting}
We provide a synonym for the \K{CaseShow (Equal a b) a b} instance, which we call \K{CaseRecShow}, as before with \K{ProofEq}; likewise a synonym for the matching function:
\begin{lstlisting}
caseShow :: ∀a b. CaseRecShow a b => b -> String
caseShow = caseShow' @(Equal a b) @a
\end{lstlisting}

@@

We can continue along the lines of the previous examples to consider the approach separately from concrete ones. One can use the technique, following the abstracted pattern below:

\begin{lstlisting}
class DispatchRec_Aux (p :: Bool)
                    (a_1 :: k_1) ... (a_n :: k_n) (b :: *) where
  dispatchRec_Aux :: X b

@@

dispatchRec :: ∀a_1 ... a_n b. DispatchRec a_1 ... a_n b => X b
dispatchRec = dispatchRec_Aux @(P a_1 ... a_k b) @a_1 ... @a_k ... @a_n
\end{lstlisting}

@@

For consumer functions, such as \K{subterms} and \K{gshow}, the type \K{X b} typically is instantiated with \K{b -> X} for some datatype \K{X}. The pattern fits equally well for defining producers, such as \K{garbitrary} from the \textsf{basic-sop} package, where \K{X b} would be instantiated with \K{Gen b}.
